Stock enhancement undermines community stability

### 2022-01-31 rmd/manuscript.Rmd
edit: manuscript

**Affiliations:** ^1^ Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA ^2^ XXX ^3^ Faculty of Environmental Earth Science, Hokkaido University, Sapporo, Hokkaido 060-0810, Japan ^4^ XXX

@@

Although stock enhancement with captive-bred organisms entails ecological risks, this method remains to be a common practice in conservation and fisheries. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks. Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Here, we use a simple theory to show that stock enhancement undermines long-term community stability by disrupting the sensitive balance of species interactions in natural communities. Stock enhancement destabilized community dynamics through facilitating competitive exclusion across a wide range of simulated scenarios. The present study further provided empirical support for this theoretical prediction using long-term data of Japanese stream fish communities. As our theory predicts, stream fish communities showed greater temporal fluctuations and reduced biodiversity in watersheds with intensive stocking of hatchery masu salmon - one of the most valuable freshwater resources in the region. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

@@

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to the sustainable management of natural resources [@cardinaleBiodiversityLossIts2012]. Captive breeding is a common practice to conserve or enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996]. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles) [@arakiGeneticEffectsCaptive2007a; @krkosekDecliningWildSalmon2007], this method is still pervasive in conservation [@snyderLimitationsCaptiveBreeding1996; @griffithsCaptiveBreedingReintroduction2008; @haagNorthAmericanFreshwater2012] and fisheries [@kitadaEconomicEcologicalGenetic2018; @arakiHatcheryStockingHelp2010]. In fisheries for example, billions of hatchery-reared individuals are released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species through density-dependent feedback underpins emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999a; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Stock enhancement may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@weberInteractionsHatcheryWild2003; @careyNativeInvadersChallenges2012]. Hence, this form of species management may have significant consequences for community dynamics in the long term. Evidence for this hypothesis is lacking, however.

Here, we use a simple theory to show that stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999a]. Our theory illuminates that stock enhancement interferes with the internal regulation process of self-sustainable communities, compromising the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by showing its congruence with Japanese stream fish communities, where \>10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes [@kitadaLessonsJapanMarine2020]. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

@@

$$
N_{i,t+1} = (N_{i,t}+\phi_i R_t)~exp[r_i(1-\frac{N_{i,t}+\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i})]~exp(\epsilon_{i,t})
$$

@@

$$
\phi_i = \left\{
\begin{array}{}
f_R & (i = 1)\\
0 & (i \ne 1)
\end{array}
\right.
$$

@@

The past few decades have witnessed substantial efforts to unveil ecological consequences of captive breeding, and there is a general agreement that the distinctive features of captive-bred individuals bring fitness disadvantages in the wild [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996]. As such, numerous studies suggested that stock enhancement with captive breeding must be carefully employed given the potential negative impacts on wild populations [@christieGeneticAdaptationCaptivity2012; @arakiGeneticEffectsCaptive2007a]. Despite the significant attention to the fate of captive-bred individuals, current schemes rarely consider the self-regulation process of biodiversity. The results of this study suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community. While our empirical example is limited to stream fish communities, the consistent emergence of destabilizing effect across simulated ecological scenarios implies that this ecological phenomenon is pervasive across terrestrial and aquatic taxa.

>>>>

**Affiliations:** ^1^ Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA ^2^ Salmon and Freshwater Fisheries Research Institute, Hokkaido Research Organization, Eniwa, Hokkaido 061-1433, Japan ^3^ Faculty of Environmental Earth Science, Hokkaido University, Sapporo, Hokkaido 060-0810, Japan ^4^ XXX

@@

There is increasing awareness that stock enhancement with captive-bred organisms brings ecological risks to wild populations. Still, we may underestimate the negative impact of stock enhancement because current schemes rarely consider the consequences for the whole ecological community. Here, we use theory and long-term data of Japanese stream fish communities to show that stock enhancement undermines community stability. Our theory predicted that stock enhancement destabilizes community dynamics by facilitating competitive exclusion. Consistent with this prediction, fish communities showed greater temporal fluctuations and reduced biodiversity in rivers with the intensive stocking of hatchery masu salmon - one of the most valuable freshwater resources in the region. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

@@

Human demands for natural resources are ever increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@cardinaleBiodiversityLossIts2012]. Captive breeding is a common practice to enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996]. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles) [@arakiGeneticEffectsCaptive2007; @krkosekDecliningWildSalmon2007], this method is still pervasive in conservation [@snyderLimitationsCaptiveBreeding1996; @griffithsCaptiveBreedingReintroduction2008; @haagNorthAmericanFreshwater2012] and natural resource management [@kitadaEconomicEcologicalGenetic2018; @arakiHatcheryStockingHelp2010]. In fisheries for example, billions of hatchery-reared individuals are released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species through density-dependent feedback underpins emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999a; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Stock enhancement may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@weberInteractionsHatcheryWild2003; @careyNativeInvadersChallenges2012]. Hence, this form of species management may undermine the ecological process that allows competing species to coexist, ultimately compromising the long-term community stability. Evidence for this hypothesis is lacking, however.

Here, we use a simple theory to show that stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999a]. Our theory illuminates that stock enhancement interferes with the internal regulation process of self-sustainable communities, compromising the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by showing its congruence with Japanese stream fish communities, where \>10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes [@kitadaLessonsJapanMarine2020].

@@

$$
N_{i,t+1} = (N_{i,t}+\phi_i R_t)~exp\left[r_i\left(1-\frac{N_{i,t}+\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i}\right)\right]~exp(\epsilon_{i,t})
$$

@@

$$
\phi_i = \left\{
\begin{array}{cc}
f_R & (i = 1)\\
0 & (i \ne 1)
\end{array}
\right.
$$

@@

The past few decades have witnessed substantial efforts to unveil ecological consequences of captive breeding, and there is a general agreement that the distinctive features of captive-bred individuals bring fitness disadvantages in the wild [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996]. As such, numerous studies suggested that stock enhancement with captive breeding must be carefully employed given the potential negative impacts on wild populations [@christieGeneticAdaptationCaptivity2012; @arakiGeneticEffectsCaptive2007]. Despite the significant attention to the fate of captive-bred individuals, current schemes rarely consider the self-regulation process of biodiversity. The results of this study suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community. While our empirical example is limited to stream fish communities, the consistent emergence of destabilizing effect across simulated ecological scenarios implies that this ecological phenomenon is pervasive across terrestrial and aquatic taxa.
